\documentclass[12pt]{report}
\usepackage[a4paper, total={6.5in, 8in}]{geometry}
\usepackage{xargs}
\usepackage{amsmath,amssymb}
\usepackage{hyperref}
\usepackage{physics}
\usepackage{graphicx}

\newcommand{\sgn}{\text{sgn}}
\newcommand{\seq}[1]{\langle #1\rangle}
\newcommand{\asc}[1]{\upharpoonleft #1 \upharpoonright}
\newcommand{\hc}{^\dagger}
\newcommand{\inv}{^{-1}}
\newcommand{\Sym}{\text{Sym}}
\newcommand{\normord}[1]{:\mathrel{\mspace{2mu}#1\mspace{2mu}}:}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}

\begin{document}

	\title{Bosonic and fermionic particle-preserving coherent states, their dynamics and applications\\\hfill\\Transfer Report}
	\author{Michal Horanský}
	\maketitle
	
	\chapter*{Abstract}
	Coherent states-based methods have been previously developed and utilised for particle-preserving bosonic systems, such as the Bose-Hubbard model. This work is divided into two parts: Firstly, a new method is proposed for the bosonic systems, where the basis coherent states are propagated separately to create a frozen dynamical basis, which lowers the time-complexity significantly. Second, an analogous method is proposed for particle-preserving fermionic systems, and particularly the electronic structure of a molecule, for which the goal of this method is to find the ground state. For this, multiple methods are outlined, and relevant properties of the fermionic coherent states are obtained.
	
	
	\tableofcontents
	
	
	\chapter{Background}
	
	To find the dynamics of a given system, one typically needs to solve the Schrodinger equation given by the system's Hamiltonian. For complex systems which can be found in nature, an analytic solution rarely exists. To solve this partial differential equation computationally, one needs to chose a basis for the Hilbert space and, since it typically is infinite or even continuous, sample it to obtain a basis sample onto which the initial state is decomposed and on which it is propagated. This sampling choice is always limiting and must be carefully justified, as omitting elements of the Hilbert space may render the analysis inaccurate.
	
	Coherent states are particular states in the Hilbert space which follow classical trajectories, remain coherent, and minimise position-momentum uncertainty. Choosing them as the basis, a very small sample is typically suitable to capture the physics of a wide variety of systems and starting conditions \cite{grossmann}. This is an example of a semi-classical approximation.
	
	\include{sections/1_semi_classical_methods}
	
	
	
	\include{sections/2_history_of_coherent_states}
	
	
	
	\include{sections/3_mathematical_approach_to_coherent_states}
	
	
	
	
	
	\chapter{Current work}
	In this chapter I outline the application of the mathematical construction of $SU(M)$ coherent states in particular to bosonic systems (such as the ones described in Sec. \ref{sec:bosonic1}, \ref{sec:bosonic2}) and fermionic systems (such as the one described in Sec. \ref{sec:fermionic1}).
	
	\include{sections/4_bosonic_SU_M_coherent_states}
	
	
	
	\include{sections/5_fermionic_SU_M_coherent_states}
	
	
	
	\chapter{Outlook}
	
	In this chapter the scope of work which has been done is summarised and an outline of the next steps of the research is presented.
	
	\section{Summary of original work done}
	
	During the first year of this doctorate degree, I have done the following original work:
	\begin{itemize}
	\item I have generalised the matrix form of the fully-variational equations of motion on a discrete CS sample as quoted by Qiao and Grossmann \cite[App. B]{grossmann}, which is also applicable to fermionic coherent states.
	\item Based on the suggestion by my supervisor Prof Dmitry Shalashilin, I have built a framework in Python 3 for bosonic $SU(M)$ dynamics on an arbitrary second-order Hamiltonian for both the fully variational method and the new separate-basis-trajectory method, and implemented the Bose-Hubbard and the displaced harmonic trap systems in this framework.
	\item Using this framework, I have achieved convergence for two- and three-mode systems for the Bose-Hubbard model on a scale which matches that of the work of Qiao and Grossmann, demonstrating the utility of the separate-basis-trajectory method. I have also extended the scope of this investigation onto problems with initial wavefunctions which are not pure bosonic coherent states.
	\item I have derived the general matrix element of a product sequence of fermionic creation and annihilation operators for fermionic $SU(M)$ coherent states, which allows the evaluation of matrix elements of arbitrary particle-preserving operators, including the Hamiltonian.
	\item I have built the skeleton of a framework in Python 3 using PySCF to translate the single-electron and electron-exchange energy integrals for an arbitrary molecule into a mode system onto which fermionic $SU(M)$ coherent state-based methods may be applied.
	\end{itemize}
	
	\section{Research plan for the duration of the doctorate programme}
	I propose the following tasks for the remaining duration of my doctorate programme to be done within this research project:
	\begin{itemize}
	\item Finishing the bosonic dynamics for separate basis trajectories (estimated duration: 2 months). This entails
		\begin{itemize}
		\item Implementing arbitrary initial wavefunction sampling techniques and testing various conditioning methods to maximise numerical stability
		\item Scaling the mode and particle numbers to see the limits of this method
		\item For high mode and particle numbers, it is no longer possible to calculate the "true" solution on the full occupancy basis; comparison data must be collected by other techniques, e.g. MCSCF.
		\end{itemize}
	\item Obtaining a succesful estimate for the ground state of a complex molecule using $SU(M)$ fermionic coherent states (estimated duration: 2 months).
	\item Further investigation of the mathematical properties of $SU(M)$ fermionic coherent states, with the focus on lowering the time complexity of calculating the Hamiltonian matrix elements (estimated duration: 3 months).
	\item Benchmarking different methods for obtaining the molecular ground state (estimated duration: 3 months).
	\item Investigating alternative constructions for obtaining the molecular electronic ground state, such as the boson-analogous construction recently proposed by Prof Dmitry Shalashilin (estimated duration: 6 months).
	\item Applying the fermionic $SU(M)$ methods for dynamics to other particle-preserving fermionic systems (e.g. qubit systems, electron gas with high chemical potential etc) (estimated duration: 6 months).
	\item Investigating the underlying connection between bosonic and fermionic $SU(M)$ coherent states and generalising it to other dynamical groups, which massively enlargens the scale of the proposed methods in terms of fermionic physical systems which may be analysed (estimated duration: 1 year in parallel with other work)
	\end{itemize}
	
	\section{Data management plan}
	The project is purely theoretical/computational, and thus no collection of laboratory data is required.
	
	Collection of numerical data from the frameworks developed by me has so far been done on my personal computer and stored on a github repository. As I move ahead to upscale the data collection, especially for large-mode bosonic systems and complex molecules, I will create a designated workflow for collecting data on the MacBook machine provided to me by the Faculty, so that long-term data collection may be put into place without interrupting the development workflow.
	
	In case the computational complexity of the largest systems we choose to analyse exceeds what may reasonably covered by a single machine, parallelisation or employing high-performance computing systems will be considered; however, given that the utility we are trying to demonstrate is that of a low-time-complexity approach, this will probably not be necessary.
	
	
	
	%\clearpage
	\begin{thebibliography}{10}

	\addcontentsline{toc}{chapter}{Bibliography}
	
	\bibitem{curse_of_dimensionality}
	Shalashilin, D. V. (2011), Multiconfigurational Ehrenfest approach to quantum coherent dynamics in large molecular systems. \textit{Faraday Discussions}, \textbf{153}, pp. 105--116
	
	\bibitem{semiclassical_generic_1}
	Crespo-Otero, R., Barbatti, M. (2018), Recent Advances and Perspectives on Nonadiabatic Mixed Quantum–Classical Dynamics. \textit{Chem. Rev.}, \textbf{118}, pp. 7026--7068
	
	\bibitem{semiclassical_generic_2}
	Mattos, R. S., Mukherjee, S., Barbatti, M. (2024), Quantum Dynamics from Classical Trajectories. \textit{J. Chem. Theory Comput.}, \textbf{20}, pp. 7728--7743
	
	\bibitem{mulliken}
	Mulliken, R. S. (1932), Electronic Structures of Polyatomic Molecules and Valence. II. General Considerations. \textit{Phys. Rev.}, \textbf{14}, pp. 49--71
	
	\bibitem{harmonic_classical_states}
	Schrodinger, E. (1926), Der stetige übergang von der Mikro-zur Makromechanik. \textit{Naturwiss}. 14, pp. 664-666. Translated into English in Schrodinger, E. (1928), \textit{Collected Papers in Wave Mechanics}. 1st edn. London: Blackie \& Son, pp. 41--44
	
	\bibitem{field_coherent_states}
	Glauber, R. J. (1963), Coherent and Incoherent States of the Radiation Field. \textit{Phys. Rev.}, \textbf{131}, 6, pp. 2766--2788
	
	\bibitem{no_unity}
	Klauder, J. R., Skagerstam, B. S. (1985), \textit{Coherent States: Applications in Physics and Mathematical Physics}. 1st eng. edn. Singapore: World Scientific.
	
	\bibitem{perelomov_og}
	Perelomov, A. M. (1972), Coherent states for arbitrary Lie group. \textit{Commun. Math. Phys.}, \textbf{26}, pp. 222--236
	
	\bibitem{gilmore_og}
	Gilmore, R. (1972), Geometry of symmetrized states. \textit{Ann. Phys.}, \textbf{74}, 2, pp. 391--463
	
	\bibitem{ZFG}
	Zhang, W. M., Feng, D. H., Gilmore, R. (1990), Coherent states: Theory and some applications. \textit{Rev. Mod. Phys.}, \textbf{62}, pp. 867--927
	
	\bibitem{aguiar}
	Viscondi, T. F., Grigolo, A., de Aguiar, M. A. M. (2015), Semiclassical Propagator in the Generalized Coherent-State Representation. \href{https://doi.org/10.48550/arXiv.1510.05952}{arXiv:1510.05952 \textbf{[quant-ph]}}
	
	\bibitem{optical_lattices}
	Block, I., Dalibard, J., Zwerger, W. (2008), Many-body physics with ultracold gases. \textit{Rev. Mod. Phys.}, \textbf{80}, pp. 885--964
	
	\bibitem{grossmann}
	Qiao, Y., Grossmann, F. (2021), Exact variational dynamics of the multimode Bose-Hubbard model based on $SU(M)$ coherent states. \textit{Phys. Rev. A}, \textbf{103}, 042209
	
	\bibitem{green}
	Green, J. A., Shalashilin, D. V. (2019), Simulation of the quantum dynamics of indistinguishable bosons with the method of coupled coherent states. \textit{Phys. Rev. A}, \textbf{100}, 013607
	
	\bibitem{gellmann}
	Bertlmann, R. A., Krammer, P. (2008), Bloch vectors for qudits. \href{ 	
https://doi.org/10.48550/arXiv.0806.1174}{arXiv:0806.1174 \textbf{[quant-ph]}}

	\bibitem{buonsante}
	Buonsante, P., Penna, V.(2008), Some remarks on the coherent-state variational approach to nonlinear boson models. \textit{J. Phys. A: Math. Theor.}, \textbf{41}, 175301
	
	\bibitem{sampling_algorithm}
	Grigolo, A., Viscondi, T. F., de Aguiar, M. A. M. (2016), Multiconfigurational quantum propagation with trajectory-guided generalized coherent states. \textit{J. Chem. Phys.}, \textbf{144}, 094106
	
	\bibitem{imaginary_timeprop}
	Pederiva, F., Roggero, A., Schmidt, K. E. (2017), "Variational and Diffusion Monte Carlo Approaches to the Nuclear Few- and Many-Body Problem" in \textit{An Advanced Course in Computational Nuclear Physics: Bridging the Scales from Quarks to Neutron Stars.} Springer International Publishing, Cham, pp. 401--476
	
	\bibitem{charpol}
	Horn, R. A., Johnson, C. R. (2012), \textit{Matrix Analysis}. 2nd edn. Cambridge University Press, Cambridge. \href{https://doi.org/10.1017/CBO9781139020411}{doi.org/10.1017/CBO9781139020411}
	
	\bibitem{modified_cauchy_binet}
	Chapman, A., Miyake, A. (2018), Classical simulation of quantum circuits by dynamical localization: analytic results for Pauli-observable scrambling in time-dependent disorder. Available at \href{https://arxiv.org/pdf/1704.04405v2}{arxiv.org/pdf/1704.04405v2 [quant-ph]}

	\end{thebibliography}
	\bibliographystyle{unsrt}
	
	\appendix
	
	\include{sections/A_auxilary_theorems}
	
	
	
	
	

	
	
\end{document}